---
title: "Codebook"
author: "Shishir"
date: "27/08/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
The `run_analysis.R` script performs the data preparation and then followed by the 5 steps required as described in the course project’s definition.

**1.Download the dataset**
  Dataset downloaded and extracted under the folder called `UCI HAR Dataset`.

**2. Assign each data to variables**
`features` <- `features.txt` : 561 rows, 2 columns
The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ.
`activities` <- `activity_labels.txt` : 6 rows, 2 columns
List of activities performed when the corresponding measurements were taken and its codes (labels)
`subject_test` <- `test/subject_test.txt` : 2947 rows, 1 column
contains test data of 9/30 volunteer test subjects being observed
`x_test` <- `test/X_test.txt` : 2947 rows, 561 columns
contains recorded features test data
`y_test` <- `test/y_test.txt` : 2947 rows, 1 columns
contains test data of activities’code labels
`subject_train` <- `test/subject_train.txt` : 7352 rows, 1 column
contains train data of 21/30 volunteer subjects being observed
`x_train` <- `test/X_train.txt` : 7352 rows, 561 columns
contains recorded features train data
`y_train` <- `test/y_train.txt` : 7352 rows, 1 columns
contains train data of activities’code labels

**Merges the training and the test sets to create one data set**
`X` (10299 rows, 561 columns) is created by merging x_train and x_test using rbind() function
`Y `(10299 rows, 1 column) is created by merging y_train and y_test using rbind() function
`Subject` (10299 rows, 1 column) is created by merging subject_train and subject_test using rbind() function
`Merged_Data` (10299 rows, 563 column) is created by merging Subject, Y and X using cbind() function

**Extracts only the measurements on the mean and standard deviation for each measurement**
`TidyData `(10299 rows, 88 columns) is created by subsetting Merged_Data, selecting only columns: subject, code and the measurements on the mean and standard deviation (std) for each measurement

**Uses descriptive activity names to name the activities in the data set**
Entire numbers in `code` column of the TidyData replaced with corresponding activity taken from second column of the `activities` variable

**Appropriately labels the data set with descriptive variable names**
`code` column in `TidyData` renamed into `activities`
All `Acc` in column’s name replaced by `Accelerometer`
All `Gyro` in column’s name replaced by `Gyroscope`
All `BodyBody` in column’s name replaced by `Body`
All `Mag` in column’s name replaced by `Magnitude`
All start with character `f` in column’s name replaced by` Frequency`
All start with character `t` in column’s name replaced by `Time`

**From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject**
`FinalData` (180 rows, 88 columns) is created by sumarizing `TidyData` taking the means of each variable for each activity and each subject, after groupped by subject and activity.
Export `FinalData` into `FinalData.txt` file.


```{r eval=FALSE, message=TRUE, include=FALSE, paged.print=TRUE}
#package
library(dplyr)
filename <- "Coursera_DS3_Final.zip"

# Checking if archieve already exists.
if (!file.exists(filename)){
  fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  download.file(fileURL, filename, method="curl")
}  

# Checking if folder exists
if (!file.exists("UCI HAR Dataset")) { 
  unzip(filename) 
}

#datas
features <- read.table("UCI HAR Dataset/features.txt", col.names = c("n","functions"))
activities <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x_test <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x_train <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")

#Merges the training and the test sets
X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
Merged_Data <- cbind(Subject, Y, X)

#Extracts only the measurements on the mean and standard deviation for each measurement.
TidyData <- Merged_Data %>% select(subject, code, contains("mean"), contains("std"))

#activity names
TidyData$code <- activities[TidyData$code, 2]

#labeling
names(TidyData)[2] = "activity"
names(TidyData)<-gsub("Acc", "Accelerometer", names(TidyData))
names(TidyData)<-gsub("Gyro", "Gyroscope", names(TidyData))
names(TidyData)<-gsub("BodyBody", "Body", names(TidyData))
names(TidyData)<-gsub("Mag", "Magnitude", names(TidyData))
names(TidyData)<-gsub("^t", "Time", names(TidyData))
names(TidyData)<-gsub("^f", "Frequency", names(TidyData))
names(TidyData)<-gsub("tBody", "TimeBody", names(TidyData))
names(TidyData)<-gsub("-mean()", "Mean", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-std()", "STD", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-freq()", "Frequency", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("angle", "Angle", names(TidyData))
names(TidyData)<-gsub("gravity", "Gravity", names(TidyData))

#average of each variable for each activity and each subject.
FinalData <- TidyData %>%
  group_by(subject, activity) %>%
  summarise_all(funs(mean))
write.table(FinalData, "FinalData.txt", row.name=FALSE)

```


